NMRlipids IV: Headgroup \& glycerol backbone structures, and cation binding in bilayers with PS lipids

### 2017-12-15 Manuscript/manuscriptPS.tex
Include the recent contributions by Markus Miettinen to the version containing only PS results

\subsection{PS headgroup and glycerol backbone structure in simulations and experiments}
The headgroup order parameters of DOPS and POPS bilayers 
from different experiments and simulations are shown in
Figs. \ref{HGorderParametersPS}.
None of the tested models gives satisfactory agreement with
experiments for order parameters in headgroup $\alpha$ and $\beta$
carbons.
\begin{figure}[]
  \centering
  \includegraphics[width=9.0cm]{../Figs/HGorderparametersPS.eps}
  \caption{\label{HGorderParametersPS}
    Order parameters for DOPS headgroup and glycerol
    backbone from simulations with different models and experiments without CaCl$_2$ 
    Experimental data from \cite{browning80} contains 0.1M of NaCl.
    Signs are taken from experiments for POPS described in Supplementary Information.
  }
  \todo{Check and report all the counterions.}
  \todo{Glycerol backbone order parameters should be available from the spectra measured by Tiago Ferreira.}
\end{figure}

@@

>>>>

\subsection{PS headgroup and glycerol backbone structure in simulations and experiments}
The headgroup order parameters of DOPS and POPS bilayers 
from different experiments and simulations are shown in
Figs. \ref{HGorderParametersPS}.
None of the tested models gives satisfactory agreement with
experiments for order parameters in headgroup $\alpha$ and $\beta$
carbons.
\begin{figure}[]
  \centering
  \includegraphics[width=9.0cm]{../Figs/HGorderparametersPS.pdf}
  \caption{\label{HGorderParametersPS}
    Order parameters for PS headgroup and glycerol
    backbone from simulations with different models and experiments without CaCl$_2$ 
    Experimental data from \cite{browning80} contains 0.1M of NaCl.
    Signs are taken from experiments for POPS described in Supplementary Information.
    The vertical bars shown for the computational values are not error bars, but demonstrate that for these systems we had at least two data sets; the ends of the bars mark the extreme values from the sets, and the dot marks their measurement-time-weighted average.
  }
  \todo{Check and report all the counterions.}
\end{figure}

\begin{figure}[]
  \centering
  \includegraphics[width=9.0cm]{../Figs/comparisonTablePS.pdf}
  \caption{\label{comparisonTablePS}
  Rough subjective ranking of force fields based on Figure~\ref{HGorderParametersPS}. Here ÒMÓ indicates a magnitude problem, ÒFÓ a forking problem; letter size increases with problem severity. Color scheme: Òwithin experimental errorÓ (dark green), Òalmost within experimental errorÓ (light green), Òclear deviation from experimentsÓ (light red), and Òmajor deviation from experimentsÓ (dark red). The $\Sigma$-column shows the total deviation of the force field, when individual carbons are given weights of 0 (matches experiment), 1, 2, and 4 (major deviation). For full details of the assessment, see Supplementary Information.
  }
\end{figure}

@@

\pagebreak
\section{Details of the rough subjective force field ranking (Fig.~\ref{comparisonTablePS})}

The assessment was based fully on the Fig.~\ref{HGorderParametersPS}.
%
First, for each carbon (the columns in Fig.~\ref{HGorderParametersPS}) in each force field (the rows),
we looked separately at deviations in magnitude and forking.

{\bf Magnitude} deviations, i.e., how close to the experimentally obtained C--H order parameters (OPs)
the force-field-produced OPs were.
%
For each carbon, the following 5-step scale was used:
%
\begin{description}
\item [0 (~):] \noindent {More than half of all the calculated OPs (that is, of all different hydrogens in all different lipids) were within the {\it subjective sweet spots} (SSP, blue-shaded areas in Fig.~\ref{HGorderParametersPS}).}
%
\item [1 ({\textsf{\tiny M}}):] \noindent {All the calculated OPs  were $< 0.03$ units away from the SSP.}
%
\item [2  ({\textsf{\small M}}):] \noindent {All the calculated OPs  were $< 0.05$ units away from the SSP.}
%
\item [3 ({\textsf{\large M}}):] \noindent {All the calculated OPs  were $< 0.10$ units away from the SSP.}
%
\item [4 ({\textsf{\Large M}}):] \noindent {Some of the calculated OPs  were $> 0.10$ units  away from the SSP.}
\end{description}

{\bf Forking} deviations, i.e., how well the difference in order parameters of two hydrogens attached to a given carbon matched that obtained experimentally. Note that this is not relevant for $\beta$ and $\mathrm{g_2}$, which have only one hydrogen. For the $\alpha$ carbon,  for which a considerable forking of 0.105 is experimentally seen, the following 5-step scale was used:
\begin{description}
\item [0 (~):] \noindent {The distance $D$ between the dots (that mark the measurement-time-weighted averages in Fig.~\ref{HGorderParametersPS}) was $0.08 < D< 0.13$ units for all the calculated OPs (that is, for all different lipids).}
%
\item [1 ({\textsf{\tiny F}}):] \noindent {$(0.06 < D < 0.08)$ OR $(0.13 < D < 0.15)$.}
%
\item [2  ({\textsf{\small F}}):] \noindent {$(0.04 < D < 0.06)$ OR $(0.15 < D < 0.17)$.}
%
\item [3 ({\textsf{\large F}}):] \noindent {$(0.02 < D < 0.04)$ OR $(0.17 < D < 0.19)$.}
%
\item [4 ({\textsf{\Large F}}):] \noindent {$(D<0.02)$ OR $(0.19<D)$.}
\end{description}
%
For the $\mathrm{g_3}$ carbon, for which no forking is indicated by experiments, the following 5-step scale was used:
%
\begin{description}
\item [0 (~):] \noindent {$ D< 0.02$.}
%
\item [1 ({\textsf{\tiny F}}):] \noindent {$0.02 < D < 0.04$.}
%
\item [2  ({\textsf{\small F}}):] \noindent {$0.04 < D < 0.06$.}
%
\item [3 ({\textsf{\large F}}):] \noindent {$0.06 < D < 0.08$.}
%
\item [4 ({\textsf{\Large F}}):] \noindent {$0.08 < D$.}
\end{description}
%
For the $\mathrm{g_1}$ carbon, for which a considerable forking of 0.13 is experimentally seen, the following 5-step scale was used:
%
\begin{description}
\item [0 (~):] \noindent {$0.11 < D < 0.15$.}
%
\item [1 ({\textsf{\tiny F}}):] \noindent {$(0.09 < D < 0.11)$ OR $(0.15 < D < 0.17)$.}
%
\item [2  ({\textsf{\small F}}):] \noindent {$(0.07 < D < 0.09)$ OR $(0.17 < D < 0.19)$.}
%
\item [3 ({\textsf{\large F}}):] \noindent {$(0.05 < D < 0.07)$ OR $(0.19 < D < 0.21)$.}
%
\item [4 ({\textsf{\Large F}}):] \noindent {$(D<0.05)$ OR $(0.21<D)$.}
\end{description}

Based on these assessments of magnitude and forking deviations,
each carbon was then assigned to one of the following groups:
"within experimental error"
(magnitude and forking deviations both on step 0 of the scales described above),
"almost within experimental error"
(sum of the magnitude and forking deviation steps 1 or 2),
"clear deviation from experiments"
(sum of magnitude and forking deviation steps from 3 to 5), and
"major deviation from experiments"
(sum of magnitude and forking deviation steps from 6 to 8).
These groups are indicated by colors in Fig.~4.
(Note that for $\beta$ and $\mathrm{g_2}$, for which there can be no forking,
the corresponding group assigment limits were: 0, 1, 2, and 3.)

Finally, the total ability of the force field to describe the headgroup and
glycerol structure was estimated.
To this end, the groups were given the following weights:
0 (within experimental error),
1 (almost within experimental error),
2 (clear deviation from experiments),
4 (major deviation from experiments),
and the weights of the five carbons were summed up.
The sum, given in the $\Sigma$-column of Fig.~\ref{HGorderParametersPS},
was then used to (roughly and subjectively, as should be clear from the
above description) rank the force fields.

\newpage
